---
title: "Data Science Group Project"
output:
  pdf_document: default
  html_notebook: default
---

### Required Packages
```{r}
library(e1071)
```


## Data Processing

For the 2015 data, the following five days were removed since they did not have recorded visibility data: Aug 9, Aug, 10, Aug 11, Sep 20, and Sep 21.
The table shows the number of times each average visibility values appears in the dataset.

```{r}
austin_weather_2015 <- read.csv(file = 'austin_weather_2015.csv', header = TRUE)
table(austin_weather_2015$VisibilityAvgMiles)
```

```{r}
#par(mfrow=c(2, 1))
#dev.new(width=1, height=1)
plot(austin_weather_2015$VisibilityAvgMiles~austin_weather_2015$Date, ylim = c(0, 10))
```

### Reconfigure the data

#### Generate the 2 classifications
```{r}
austin_weather_2015$Date <- NULL
austin_weather_2015$Visibility <- ifelse(austin_weather_2015$VisibilityAvgMiles > 5, 1, 0)
austin_weather_2015$Visibility <- as.factor(austin_weather_2015$Visibility)
austin_weather_2015$VisibilityAvgMiles <- NULL
table(austin_weather_2015$Visibility)
```

#### Split Into Training and Testing Data
```{r}
n <- nrow(austin_weather_2015)
set.seed(1)
train <- sample(1:n, 0.8*n)
train_austin_weather <- austin_weather_2015[train,]
test_austin_weather <- austin_weather_2015[-train,]
```

### Generating the SVM Models

#### Linear SVM 
```{r echo=TRUE}
set.seed(1)
svm.lin.2015 <- tune(svm,
                Visibility~.,
                data = train_austin_weather,
                kernel = "linear",
                ranges = list(cost = seq(0.1, 10, by = 0.1)),
                scale = T)
```

```{r include=FALSE}
print(svm.lin.2015)
print(svm.lin.2015$best.parameters) # Optimal cost value
print(svm.lin.2015$best.performance) # Optimal model's cross-validation error (testing error estimate)

svm.lin.2015.pred <- predict(svm.lin.2015$best.model)
mean(svm.lin.2015.pred != train_austin_weather$Visibility) # Optimal model's training error
```

The linear kernel's best parameters are `r svm.lin.2015$best.parameters`.
It has a training error of `r round(mean(svm.lin.2015.pred != train_austin_weather$Visibility), 5)`, a testing error of `r round(mean(svm.lin.2015.pred != test_austin_weather$Visibility), 5)`, and a 10-fold cross validation error of `r round(svm.lin.2015$best.performance, 5)`.


#### Polynomial SVM
```{r echo=TRUE}
set.seed(1)
svm.poly.2015 <- tune(svm,
                Visibility~.,
                data = train_austin_weather,
                kernel = "polynomial",
                ranges = list(cost = seq(0.001, 0.1, by = 0.001),
                              degree = seq(2, 20, by = 1)),
                scale = T)
```

```{r include=FALSE}
print(svm.poly.2015)
print(svm.poly.2015$best.parameters) # Optimal cost value
print(svm.poly.2015$best.performance) # Optimal model's cross-validation error (testing error estimate)

svm.poly.2015.pred <- predict(svm.poly.2015$best.model)
mean(svm.poly.2015.pred != train_austin_weather$Visibility) # Optimal model's training error
```

The polynomial kernel's best parameters are `r svm.poly.2015$best.parameters`.
It has a training error of `r round(mean(svm.poly.2015.pred != train_austin_weather$Visibility), 5)`, a testing error of `r round(mean(svm.poly.2015.pred != test_austin_weather$Visibility), 5)`, and a 10-fold cross validation error of `r round(svm.poly.2015$best.performance, 5)`.


#### Radial SVM
```{r echo=TRUE}
set.seed(1)
svm.rad.2015 <- tune(svm,
                Visibility~.,
                data = train_austin_weather,
                kernel = "radial",
                ranges = list(cost = seq(0.1, 5, by = 0.1),
                              gamma = seq(0.01, 0.5, by = 0.01)),
                scale = T)
```

```{r include=FALSE}
print(svm.rad.2015)
print(svm.rad.2015$best.parameters) # Optimal cost value
print(svm.rad.2015$best.performance) # Optimal model's cross-validation error (testing error estimate)

### WITH THE 80/20 SPLIT
svm.rad.2015.pred <- predict(svm.rad.2015$best.model)
mean(svm.rad.2015.pred != train_austin_weather$Visibility) # Optimal model's training error
```

The radial kernel's best parameters are `r svm.rad.2015$best.parameters`.
It has a training error of `r round(mean(svm.rad.2015.pred != train_austin_weather$Visibility), 5)`, a testing error of `r round(mean(svm.rad.2015.pred != test_austin_weather$Visibility), 5)`, and a 10-fold cross validation error of `r round(svm.rad.2015$best.performance, 5)`.


### Plotting the SVM Results

### Linear Plots
```{r}
set.seed(1)
svmfit <- svm(Visibility~., data = austin_weather_2015,
              kernel = "linear",
              cost = 1.6)

plot(svmfit, austin_weather_2015, TempAvgF~DewPointAvgF)
plot(svmfit, austin_weather_2015, TempAvgF~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, TempAvgF~SeaLevelPressureAvgInches)
plot(svmfit, austin_weather_2015, TempAvgF~WindAvgMPH)
plot(svmfit, austin_weather_2015, TempAvgF~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, DewPointAvgF~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, DewPointAvgF~SeaLevelPressureAvgInches)
plot(svmfit, austin_weather_2015, DewPointAvgF~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~WindAvgMPH)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, WindAvgMPH~PrecipitationSumInches)
```

```{r echo=FALSE}
# Testing a different method:
# x <- austin_weather_2015$TempAvgF
# y <- austin_weather_2015$DewPointAvgF
# w <- austin_weather_2015$HumidityAvgPercent
# z <- austin_weather_2015$SeaLevelPressureAvgInches
# model <- svm.rad.2015$best.model
# xgrid = expand.grid(x = y, w = z) # generating grid points
# ygrid = predict(model, newdata = xgrid) # ygrid consisting of predicted Response values
# 
# Radialsvm=svm(Visibility~., data=austin_weather_2015, kernel="radial", cost=1, gamma = 0.01, scale=T)

# lets plot the non linear decision boundary
# plot(xgrid,col=as.numeric(ygrid),pch=20,cex=0.3)
# points(austin_weather_2015$Visibility~austin_weather_2015$DewPointAvgF,col=y+1,pch=19) 
# we can see that the decision boundary is non linear
```

### Polynomial Plots
```{r}
svmfit <- svm(Visibility~., data = austin_weather_2015,
              kernel = "polynomial",
              cost = 0.006, degree = 10)

plot(svmfit, austin_weather_2015, TempAvgF~DewPointAvgF)
plot(svmfit, austin_weather_2015, TempAvgF~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, TempAvgF~SeaLevelPressureAvgInches)
plot(svmfit, austin_weather_2015, TempAvgF~WindAvgMPH)
plot(svmfit, austin_weather_2015, TempAvgF~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, DewPointAvgF~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, DewPointAvgF~SeaLevelPressureAvgInches)
plot(svmfit, austin_weather_2015, DewPointAvgF~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~WindAvgMPH)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, WindAvgMPH~PrecipitationSumInches)
```


### Radial Plots
```{r}
svmfit <- svm(Visibility~., data = austin_weather_2015,
              kernel = "radial",
              cost = 1.6, gamma = 0.38)

plot(svmfit, austin_weather_2015, TempAvgF~DewPointAvgF)
plot(svmfit, austin_weather_2015, TempAvgF~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, TempAvgF~SeaLevelPressureAvgInches)
plot(svmfit, austin_weather_2015, TempAvgF~WindAvgMPH)
plot(svmfit, austin_weather_2015, TempAvgF~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, DewPointAvgF~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, DewPointAvgF~SeaLevelPressureAvgInches)
plot(svmfit, austin_weather_2015, DewPointAvgF~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~HumidityAvgPercent)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~WindAvgMPH)
plot(svmfit, austin_weather_2015, SeaLevelPressureAvgInches~PrecipitationSumInches)
plot(svmfit, austin_weather_2015, WindAvgMPH~PrecipitationSumInches)
```


